### 2020-12-21 tex/conclusions.tex
Rewording: up tp sec. 2.3.1

\section{Conclusions}\label{sec:conclusions}
We conclude the following statements:\\
\begin{enumerate}
    \item Any system (classical or quantum) out of equilibrium (including inhomogeneities) evolves to the equilibrium state without using the transport equation as long as the local equilibrium hypothesis holds.
    \item We can obtain the transport equation from the $H$-functional.
    \item The variational entropy is proportional to the Boltzmann's $H$-functional in the limit of the high energy levels.
    \item We find an expression for the entropy to quantum gases.
    \end{enumerate}

%  \begin{equation}
%      \mathcal{H}
%  \end{equation}

\clearpage