### 2021-03-10 tex/disscussion.tex

\section{Comments and remarks}\label{sec:disscussion}

The demonstration of the classical $H$-theorem usually begins by assuming that the gas,
despite being initially out of equilibrium, can be described by a
spatially homogeneous distribution function. 
Subsequently, the time-evolution of the system occurs in such a manner that $dH/dt \leq 0$. 
Therefore, this approach does not describe the evolution to equilibrium of systems with spatial inhomogeneities.
To address this issue, in this article, we
proposed a framework that may be useful to describe the time-evolution of 
initially non-homogeneous systems. To this end, we divided the
system into small cells to conceive a system wherein the local equilibrium
hypothesis is valid in each cell but in such a manner that the total system is not homogeneous. Systems
that satisfy the previous conditions will evolve towards equilibrium, and the evolution
occurs according to $d\mathcal{H}'/dt\leq0$, Eq.~(\ref{eq:cHdef})
and $d\mathcal{H}/dt\leq0$, Eq.~(\ref{eq:qHdef}), for classical and quantum gases, respectively.
Consequently, this approach can be considered an extension of the $H$-theorem
for more realistic out-of-equilibrium systems.

The classical and quantum $H$-functionals, $\mathcal{H}'$ and $\mathcal{H}$, respectively,
correctly recover the most-probable distribution
functions in out-of-equilibrium states (locally) and when the system attains the global
equilibrium state. The relaxation process of the system is described by
monotonic functions that account for deviations from the global equilibrium.

It is clear that for describing the relaxation process of a concrete system,
it is necessary to know, at least to some approximation, the specific forms of the
monotonic functions $g'_M$ and $g_{nM}$, for classical and quantum systems.
Whereas the complete analysis of these functions is beyond the scope of the present work,
some of their properties can be predicted, \textit{e.g.} they must
be consistent both with the system relaxation times and the mechanisms of 
energy transfer between cells.

An important aspect of the framework proposed in this work is related to the entropy of systems
out-of-equilibrium. Because the functionals $\mathcal{H}'$ and $\mathcal{H}$ can be related to
the entropy of dilute gases, either classical or quantum, the fact that these functionals are defined
over a system divided into cells enables their use for defining the entropy
of out-of-equilibrium systems, other than dilute gases. Specifically, and 
derived from our previous work (\textit{e.g.} \cite{bib:nicolas2020,bib:nicolas2016}), the
$\mathcal H'$ and $\mathcal H$ functionals may serve to describe the entropy,
as well as the entropy generation, occurring during the
growth of complex physical systems, such as fractals. Possibly, studying these systems might also
shed light on the explicit functional form of $g'_M$ and $g_{nM}$.

In summary, we proposed a variational procedure to demonstrate the classical and
quantum $H$-theorems, which allowed us to describe, at a mesoscopic local view
(cell-scale), the time-evolution of an out-of-equilibrium and spatially
non-homogeneous system moving towards the equilibrium condition. In principle, this
approach would permit the investigation of the transport phenomena inherent to the
equilibration process, occurring in a system with a spatially inhomogeneous
out-of-equilibrium initial condition.